---
title: "Figures for predicted indels"
author: "Vitalii Kleshchevnikov"
date: "06/07/2018"
output: 
  html_document: 
    toc: yes
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
start_time = Sys.time()

suppressPackageStartupMessages({
  library(data.table)
  library(ggplot2)
  library(cowplot)
  library(grid)
  library(Biostrings)
  library(R.utils)
})
source("./r_utils/myggpairs.R")
source("./r_utils/loadFromFARM.R")
source("./r_utils/readIndelSummaries.R") 
source("./r_utils/dominantIndel.R")
source("./r_utils/findIndelLevels.R")
source("./r_utils/Other.R")
source("./r_utils/addIndelRepl.R") 
source("./r_utils/selectCellLines.R")
source("./r_utils/predictedK562Files.R")
source("./r_utils/renameSampleGroup.R")
source("./r_utils/readDPI7OneCellLine.R") 
source("./r_utils/calcSampleScatter.R")
source("./r_utils/plotSampleScatter.R") 
source("./r_utils/mergeSampleScatter.R")
source("./r_utils/plotFreqIndels5D.R")
source("./r_utils/caclMostFreqIndel5D.R")
plot_count = 1
my.ggsave = function(plot_count, name = NULL, width = NA, height = NA,
                     units = c("in", "cm", "mm"), device = "svg"){
  if(is.null(name)) 
    name = paste0("../results/plots/", plot_count, ".", device) else
      name = paste0("../results/plots/", name, ".", device)
    dir = gsub(basename(name),"", name)
    if(!dir.exists(dir)) dir.create(dir, recursive = T)
    ggsave(name, device = device, width = width, height = height, units = units)
    plot_count + 1
} # usage plot_count = my.ggsave(plot_count)

## Find data
data.base.dir = "../data/summary_data/"
data.set.dir = "more_indel_summaries/"
data.dir = paste0(data.base.dir, data.set.dir) 
data.files = paste0(data.dir, list.files(data.dir))
```

```{r load_seq_data, message=FALSE, warning=FALSE}
# guide sequences 
seq.set.dir = "guide_target_seq/" # _normWT
seq.dir = paste0(data.base.dir, seq.set.dir) 
seq.files = paste0(seq.dir, list.files(seq.dir))
# read sequence data
seq_data = fread(seq.files, stringsAsFactors = F)
# change column names to data.table-compatible format
setnames(seq_data, colnames(seq_data),
         gsub(" ","_",colnames(seq_data)))
seq_data[, ID := gsub("_","",ID)]
Target_string = seq_data[,DNAStringSet(Target)]
names(Target_string) = seq_data$ID
Guide_string = seq_data[,DNAStringSet(Guide)]
names(Guide_string) = seq_data$ID
# find reverse complement reversed targets
Target_string[seq_data$PAM_Direction == "REVERSE"] = reverseComplement(Target_string[seq_data$PAM_Direction == "REVERSE"])
seq_data[PAM_Direction == "REVERSE", PAM_Location := nchar(Target) - PAM_Location]
seq_data[PAM_Direction == "REVERSE", PAM_Direction := "FORWARD"]
```

```{r load_all_cell_lines}
## Load and process data
####/ set some options
N_reads_filter = 20
inv = F

# define order of variables and colors for plots
source("./r_utils/order_var_script.R")

# choose thresholds for dominant indel
dominant_indel_freq = 0.2 # individual indels
####\

# select relevant files, choose a single sample group (only K562 for now):
new.DPI7.data.files = predictedFiles(data.files, sample_shared_name = "K562|_CAS9_12NA|Predicted_5000", predicted = "Predicted_5000")

# read and use gRNAs used for validation in the paper
validation_gRNAs = fread("../data/summary_data/old_new_validation_guides.txt", stringsAsFactors = F)
# change column names to data.table-compatible format
setnames(validation_gRNAs, colnames(validation_gRNAs),
         gsub(" ","_",colnames(validation_gRNAs)))
# make sure gRNAs with pam error are not in the set
pam_errors = fread("../data/summary_data/pam_errors.txt", stringsAsFactors = F)
validation_gRNAs = validation_gRNAs[!New_Oligo_Id %in% pam_errors$ID]
validation_gRNAs = validation_gRNAs[!Old_Oligo_Id %in% pam_errors$ID]


data = readIndelSummaries(new.DPI7.data.files, guide_list = validation_gRNAs$New_Oligo_Id, inv = inv)
# select guides present in all samples
data = inAllSamples(data[Total_reads > N_reads_filter],.(Oligo_Id))

# how many guides per sample?
guidesPerSample(data)
Sys.time() - start_time
```

```{r all_freq_scatterplot_repl, fig.height=7.5, fig.width=7.5}
## replace with the scatterplot function to make this plot, then generate this file with FORECasT-ver1, commit, create a copy for FORECasT-ver2, generate, commit 
data = renameSampleGroup(data)
data_repl = copy(data)
# calculate means of normalised and cumulative reads across replicates
data = addIndelRepl(data)

#sanity check
unique(data[Oligo_Id == Oligo_Id[1] & 
              Most_Common_Indel == Most_Common_Indel[1] &
              sample_group == "K562"]) # should be one line
data[, sum(sum_MCI_Reads) > sum_Total_reads, by = .(Oligo_Id, sample_group)][,mean(V1)] # should be 0

data.indels = unique(data[, .(Oligo_Id, Most_Common_Indel, mean_Norm_MCI_Reads, Norm_MCI_Reads, few_indel_classes, complex_indel_classes, sample_group)])
# set order of indel classes
data.indels[, few_indel_classes := factor(few_indel_classes,
                                                levels = few_indel_classes_levels)]
all_freq = unique(data.indels[, .(Oligo_Id, Most_Common_Indel, few_indel_classes, Norm_MCI_Reads, sample_group)])

```

## Figure S20. Comparison of allele frequencies from predictions and replicates  

Density of frequency of mutations estimated by two approaches. Left panel: predicted (y-axis) vs combined measurement (x-axis); right panel: measurement replicate 1 (x-axis)  vs measurement replicate 2 (y-axis), both categorised by mutation type (rows), with Pearson’s correlation given. I+D mutations and I>2 mutations are not predicted by the model.  

```{r, fig.height=12, fig.width=6}
data_repl2 = data_repl[sample_group == "K562" & sample_ID != "ST_Feb_2018_CAS9_12NA_1600X_DPI7"]
predicted.pairs = calcSampleScatter(data_repl2,
                                    split_by = "few_indel_classes",
                                    value.var = "Norm_MCI_Reads",
                                    x_group = "ST_June_2017_K562_800x_LV7A_DPI7",
                                    y_group = "ST_June_2017_K562_800x_LV7B_DPI7",
                                    x_name = "replicate 1", y_name = "replicate 2",
                                    data_pred = all_freq, sample_group_sel = "K562",
                                    pred_group = "Predicted",
                                    type_samples = "Replicates", type_predicted = "Predicted")
plotSampleScatter(predicted.pairs, binwidth = 0.025)
plot_count = my.ggsave(plot_count, name = "Figure_S20", width = 8, height = 12, units = "in")
```

```{r}
Sys.Date. = Sys.Date()
Sys.Date.
session_info. = devtools::session_info()
session_info.
Sys.time() - start_time
```